# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data


```{r echo = TRUE}

Sys.setlocale("LC_TIME", "en_US.UTF-8")

activityOriginal <- read.csv('activity.csv',stringsAsFactors = F);
activityDF <- activityOriginal[complete.cases(activityOriginal),];
activityDF$date <- as.Date(activityDF$date, format = "%Y-%m-%d");
perDayDF <- aggregate(activityDF$steps, by=list(date = activityDF$date), FUN=sum);
colnames(perDayDF) <- c("date","steps");

avgPerInterval <- aggregate(activityDF$steps, by=list(interval = activityDF$interval), FUN=mean);

colnames(avgPerInterval) <- c("interval","meanSteps");
maxInterval <- avgPerInterval[order(avgPerInterval$meanSteps, decreasing=TRUE), ][1,];

```


## What is mean total number of steps taken per day?

```{r echo = TRUE}
 hist(perDayDF$steps,main="Steps by Day",
       xlab="Number of steps")

```

The mean steps taken per day is:

```{r echo = TRUE}
mean(perDayDF$steps);
```

The median steps taken per day is:

```{r echo = TRUE}
median(perDayDF$steps);
```



## What is the average daily activity pattern?

```{r echo = TRUE}
 plot(avgPerInterval$interval,avgPerInterval$meanSteps,col="black", type="l",
       main="Daily Average Steps by 5 minutes interval",
       ylab="Average Steps",
       xlab="5 minutes interval")
```

The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is:

```{r echo = TRUE}
maxInterval$interval
```
with 

```{r echo = TRUE}
maxInterval$meanSteps
```
average steps

## Inputing missing values

I used the average of the interval to fill  missing steps (with NA's)

```{r echo = TRUE}

library(plyr);

missingDF <- ddply(activityOriginal[!complete.cases(activityOriginal),], .(date,interval,steps), transform, steps = round(mean(activityDF$steps[activityDF$interval == interval]), digits = 0));

missingDF$date <- as.Date(missingDF$date, format = "%Y-%m-%d");

completeDF <- rbind(missingDF,activityDF);

completeDF$steps <- as.numeric(completeDF$steps);

perDayCompleteDF <- aggregate(completeDF$steps, by=list(date = completeDF$date), FUN=sum);
colnames(perDayCompleteDF) <- c("date","steps");

hist(perDayCompleteDF$steps,main="Steps by Day (Complete DataFrame)",
       xlab="Number of steps")


```


The mean steps taken per day (Complete Data Frame) is:

```{r echo = TRUE}
mean(perDayCompleteDF$steps);
```

With NA's the mean is:

```{r echo = TRUE}
mean(perDayDF$steps);
```



The median steps taken per day (Complete Data Frame) is:


```{r echo = TRUE}
median(perDayCompleteDF$steps);
```

With NA's the mean is:

```{r echo = TRUE}
median(perDayDF$steps);
```


## Are there differences in activity patterns between weekdays and weekends?

```{r echo = TRUE}

require(ggplot2)

activityDF$wday    <- as.factor(ifelse(weekdays(activityDF$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday")) ;

weekendDF <- subset(activityDF, wday == "Weekend");
weekdayDF <- subset(activityDF, wday == "Weekday");

head(weekendDF);


avgPerIntervalWeekend <- aggregate(weekendDF$steps, by=list(interval = weekendDF$interval), FUN=mean);


colnames(avgPerIntervalWeekend) <- c("interval","meanSteps");

avgPerIntervalWeekday <- aggregate(weekdayDF$steps, by=list(interval = weekdayDF$interval), FUN=mean);
colnames(avgPerIntervalWeekday) <- c("interval","meanSteps");

par(mfrow=c(2,1))

plot(avgPerIntervalWeekend$interval ,avgPerIntervalWeekend$meanSteps ,col="blue", type="l",
       main="Daily Average Steps by 5 minutes interval in Weekends",
       ylab="Average Steps",
       xlab="5 minutes interval");

plot(avgPerIntervalWeekday$interval ,avgPerIntervalWeekday$meanSteps ,col="red", type="l",
       main="Daily Average Steps by 5 minutes interval in Weekdays",
       ylab="Average Steps",
       xlab="5 minutes interval")

```


